Analysis of the 2014 Passenger Flight Network

### 2020-02-11 paper/main.tex
more bla

%\usepackage{lipsum}

@@

All the data-setup can be acomplished by running a single script.
This script and also all other code we have written for this project can be found in a public repository \cite{repository}.

@@

When we disregard odd degress, the seemingly linear distribution on a log log scale becomes apparent. Using least square (scipy.optimize.curve\_fit), we optimize paramters $c$ and $\gamma$ Figure~\ref{fig:degree_distribution_curves}

@@

In order to enable us to do more meaningful analysis, we enrich the dataset with additional information. We chose the airport patronage and found out, that we can get this data from Wikidata. 
For 1956 of the 3139 airports from the dataset, we can find patronage data from Wikidata. 
<describe how we do it: yearly vs. monthly + scaling from other years>
In order to populate the remaining nodes with reasonable patronages as well, we tried different techniques. 
First <using available data: runways -> factor -> calc. patronage per square feet + threshold, bad data quality. Show flow estimation iteration.>
Seccond <inspired by professor: algorithm that looks at neighboring airports patronage. Describe Algorithm. Show flow estimation iteration>

@@

<Passenger flow estimation algorithm, outcomes>

>>>>

%\usepackage{lipsum}
\usepackage{listings}

@@

All the data-setup can be acomplished by running a single script.
This script and also all other code we have written for this project can be found in a public repository on GitHub\cite{repository}.

@@

When we disregard odd degress, the seemingly linear distribution on a log log scale becomes apparent. Using least square (scipy.optimize.curve\_fit), we optimize paramters $c$ and $\gamma$ as you can see in Figure~\ref{fig:degree_distribution_curves}

@@

In order to enable us to do more meaningful analysis, we enrich the dataset with additional information. We chose the airport patronage and found out, that we can get this data from Wikidata. 
For 1956 of the 3139 airports from the dataset, we can find patronage data from Wikidata. 
We need to distinguish between monthly and yearly values and we extrapolate 2014 data from close years in case no data for the year 2014. 
In order to populate the remaining nodes with reasonable patronages as well, we tried different techniques. 
First we used data about runway surfaces that we found on ourflights.org. We summed up the surfaces of runways for every airport and then calculated the average ratio between total runway surface area and patronage of airports where both data points were present, in order to predict patronage for airports without patronage data but with runway surface data based on this average ratio. For the still remaining >1000 airports without patronage we would then apply a default value that equals the avg of the smallest 1\% of airports with patronage data from Wikidata. 
The runway data turned out to be very unclean and unreliably. Water runways and other quirks were distorting the predictions.

Inspired by Prof. Ikonnikova advice during the presentation we switched to estimating the remaining airport patronages by looking at neighboring airports with patronage data.

\begin{lstlisting}[language=Python, caption=Patronage Estimation]
G = loadEntireNetwork()
pr = -1 #previous remaining
r = 0   #remaining
while r != pr:
  pr = r
  r = 0
  ep = {} #estimated patronages
  for apid in G.nodes:
    node = G.nodes[apid]
    if node["patronage"] == None:
      neighbors = G.neighbors(apid) 
      deg = G.degree(apid)
      sum = 0
      count = 0
      for nid in neighbors: 
        n = G.nodes[nid]
        npatr = n["patronage"]
        ndeg = G.degree(nid) 
        if npatr != None: 
          sum += deg * npatr / ndeg
          count += 1
      if count == 0:
        r += 1
      else:
        avg = sum / count
        ep[apid] = avg 
  for apid in ep: 
    node = G.nodes[apid]
    node["patronage"] = ep[apid]
\end{lstlisting}

This seccond approach yields much nicer estimations, as we will see after estimating the flows between airports.

The flows are estimated by choosing reasonable initial values and then applying corrections iteratively to approach a situation where the sum of the estimated flows for each airport equal its patronage value.
The average initial values for an edge between airport a and b is chosen as: \footnote{hmm, vielleicht auch nicht den avg, sondern den total flow zwischen den airports so berechenen - ich mein, das hab ich ja gemacht!}

$avgFlow_{ab} = $

$\dfrac{patronage(a) * patronage(b)}{degree(a) * degree(b)}$

Remember that we are handling a multiedge directed graph here. There may be multiple edges between two airports pointing in different directions and/or operated by different airlines.
We could assume an equal distribution of the flows among the edges. But we chose trying to be closer to reality by weighting the flow of an edge between two airports $a$ and $b$, that is operated by airline $o$, by the geometric mean of the number of edges airline $o$ operates at airport $a$ and number of edges airline $o$ operates at airport $b$:

$flow_{ab} = $

$avgFlow_{ab} * (|E_{ab}|+|E_{ba}|) * weight_{abo}$

Where $|E_{ab}|$ is the number of directed edges from $a$ to $b$ and

$weight_{abo} = $

$\dfrac{\sqrt{degree_{o}(a) * degree_{o}(b)}}{\displaystyle\sum_{l \in L \land c,d \in \{a,b\} \land c \neq d}(\sqrt{degree_{l}(c) * degree_{l}(d)})}$

The ratios we choose here will stay intact over the course of the iterations in the next part of the algorithm.

<describe iteration step>

<Show results, compare runway to neighbor based estimation>

The complete algorithm for passenger flow estimation is too long for beeing included in this document. You can find it in the repository: ./scripts/estimate-passenger-flows.py

@@
